Microxanox: an R package for simulating an aquatic \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.

### 2022-05-06 inst/manuscript/manuscript_microxanox.Rmd
Main Changes done

---
title: "Microxanox - an R package for simulating an $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch;  Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
keywords: metadata quality; data curation; archival; long term storage; R package;
abstract: |
  Ca. 100 words.
journal: "SoftwareX"
date: "`r Sys.Date()`"

@@

<!--
> Abstract Romana of paper: The role of functional diversity in ecosystem regime shifts

> Biodiversity can increase the resilience of ecosystems facing environmental change. However, we have limited understanding if this holds true for ecosystems that respond to gradual environmental change with abrupt regime shifts.

> Here we investigated if functional diversity influences the resilience of ecosystems that abruptly transition between alternative ecosystem states. As a case study, we used a mathematical model developed by Bush et al. (2017) which describes anoxic-oxic regime shifts of aquatic ecosystems mediated by cyanobacteria, sulfate-reducing bacteria, and phototrophic sulfur bacteria. We extended this model so as to include functional diversity in one, two, or all three functional groups. Specifically, we manipulated the amount of variation in environmental tolerance and in maximum growth rate, and explored how trait variation influenced anoxic-oxic regime shifts along a gradient of oxygen influx. We found that greater diversity of either cyanobacteria or sulfate-reducing bacteria increased the resilience of the ecosystem state that these groups dominated, while greater diversity of phototrophic sulfur bacteria reduced the resilience of the state they dominated. We also found that simultaneous variation in multiple functional groups reduced, reversed, or did not change the diversity effect of individual functional groups. Collectively, our findings highlight the importance of considering multiple interacting functional groups when predicting the effect of functional diversity on ecosystem resilience
 -->

# \color{red}TODO
- \color{red}**update links** 
- \color{red}**update cross references**
\pagebreak

@@

| **Nr.** | **Code metadata description**                                | **Please fill in this column**                                                 |
|:-------:|:-------------------------------------------------------------|:-------------------------------------------------------------------------------|
| C1      | Current code version                                         | v0.9.0                                                                         |
| C2      | Permanent link to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) |
| C3      | Code Ocean compute capsule                                   |                                                                                |
| C4      | Legal Code License                                           | CC BY 4.0                                                                      |
| C5      | Code versioning system used                                  | git                                                                            |
| C6      | Software code languages, tools, and services used            | R                                                                              |
| C7      | Compilation requirements, operating environments             | R (>= 4.1.0), ADD PACKAGES                                                     |
| C8      | If available Link to developer documentation/manual          |                                                                                |
| C9      | Support email for questions                                  | Rainer.Krug@uzh.ch; Rainer@krugs.de                                            |

# Motivation and significance
Many ecosystems are exposed to gradual changes of environmental variables, to which the responses are not always as gradual as the change of the environmental variable. One example where the gradual change of a single environmental variable is causing an abrupt change of the system is the switch from an aerobic to anaerobic system. This system has been investigated by @Bush2017. We wanted to take this investigation one step further, and look at the role an increased biodiversity plays in these dynamics [@REF_NEEDED]. For this purpose, we developed this package.

@@

The aims of the `microxanox` package are twofold. Firstly, the package aims at reproducing the results shown by [@Bush2017], which is accomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED). Secondly, to take these results one step further, it includes new functionality to address our research question as published in [@REF_NEEDED].

@@

The framework used when writing this package aims at reproducibility of the results. It builds on the following main considerations:

@@

The point that the results object contains all parameter needed to run the simulation, promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations much easier.

@@

The individual simulation (`run_simulation()` function) is the working horse in this package. In this function, the ODEs are solved. The function needs only one argument - an object as created by the function `new_runsim_parameter()`. One parameter of this object is the `strain_parameter` which can be created by the function `new_strain_parameter()`. For a detailed description of the parameter, their meaning and how they are created please see the User Guide which accompanies the package or is available at [User Guide](@LINK_NEEDED)

After the parameter object has been defined, it can be used in the `run_simulation()` function. The function returns an object which is identical to the parameter, except of an additional slot containing the results. This design produces a fully reproducible object as it can be used instead of a parameter object to be fed back into the `run_simulation()` function to run the simulation again from the parameter used to generate the results from.

@@

### Finding a Steady State of the model

There are two methods for finding steady states. The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the *Replication method*). The second runs only two simulations, with step-wise and slowly temporally increasing or decreasing oxygen diffusivities and recorded of state just before change to a new oxygen diffusivity (let us term this the *Temporal method*).

#### Replication Method
The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.

#### Temporal Method
The temporal method involves two simulations for a particular system configuration (parameter set). In one simulation the oxygen diffusivity is *increased* in a step-wise fashion. In the other it is *decreased* in a step-wise fashion. That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set. Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step.

This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()`.

@@

### Extracting Stability Measures
From the raw results returned by these `run_...()` functions, the stability measures can be extracted by using the function `get_stability_measures()`. These measures include non-linearity and hysteresis measures, of the response of the simulated system to environmental change.

@@

### Regime shifts during temporal environmental change

@@

### The extent of hysteresis depends on community composition

@@

### Effects of functional diversity on regime shifts

@@

<span style="color:green">{>>More details from the paper.<<}
{>>Romana will provide two or three sentences<<}</span>

@@

# Conclusions
Set out the conclusion of this original software publication.

>>>>

---
title: "Microxanox: an R package for simulating an aquatic $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch;  Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
keywords: reproducibility; regime shift; 
abstract: |
  Ca. 100 words.
journal: "SoftwareX"
date: "`r Sys.Date()`"

@@

# \textcolor{red}{TODO}
- **\textcolor{red}{update links}**
- **\textcolor{red}{update cross references}**
- **\textcolor{red}{check keywords}**
- **\textcolor{red}{add more references}**
- **\textcolor{red}{expand on Impact and Conclusion section}**
\pagebreak

@@

| **Nr.** | **Code metadata description**                                | **Please fill in this column**                                                 |  
| :-----: | :----------------------------------------------------------- | :----------------------------------------------------------------------------- |  
|   C1    | Current code version                                         | v0.9.0                                                                         |  
|   C2    | Permanent link to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) |  
|   C3    | Code Ocean compute capsule                                   |                                                                                |  
|   C4    | Legal Code License                                           | CC BY 4.0                                                                      |  
|   C5    | Code versioning system used                                  | git                                                                            |  
|   C6    | Software code languages, tools, and services used            | [R](https://cran.r-project.org/index.html)                                     |  
|   C7    | Compilation requirements, operating environments             | [R (>= 4.1.0)](https://cran.r-project.org/index.html)                          |  
|         |                                                              | magrittr                                                                       |  
|         |                                                              | tibble                                                                         |  
|         |                                                              | ggplot2                                                                        |  
|         |                                                              | patchwork                                                                      |  
|         |                                                              | grDevices                                                                      |  
|         |                                                              | stats                                                                          |  
|         |                                                              | mgcv                                                                           |  
|         |                                                              | deSolve                                                                        |  
|         |                                                              | dplyr                                                                          |  
|         |                                                              | tidyr                                                                          |  
|         |                                                              | stringr                                                                        |  
|         |                                                              | multidplyr                                                                     |  
|         |                                                              |                                                                                |  
|         |                                                              |                                                                                |  
|         |                                                              |                                                                                |  
|         |                                                              |                                                                                |  
|         |                                                              |                                                                                |  
|   C8    | If available Link to developer documentation/manual          | \textcolor{red}{TO BE ADDED}                                                   |  
|   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Rainer@krugs.de                                            |

# Motivation and significance
Many ecosystems are exposed to gradual changes of environmental variables, to which the responses are not always as gradual as the change of the environmental variable [@REF_NEEDED]. One example where the gradual change of a single environmental variable is causing an abrupt change of the system is the switch from an aerobic to anaerobic system. This system has been investigated by @Bush2017. We wanted to take this investigation one step further, and look at the role an increased biodiversity plays in these dynamics [@REF_NEEDED]. For this purpose, we developed this package.

@@

The aims of the `microxanox` package are twofold: firstly, to reproduce the results shown by [@Bush2017], which is accomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED) and secondly, to take these results one step further, to includes new functionality to address our research question as published in [@REF_NEEDED].

@@

The framework used when writing this package aims at reproducibility of the results (see [@REF_NEEDED] for examples). It builds on the following main considerations:

@@

The point that the results object contains all parameter needed to run the simulation, promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations much easier, as can be seen in our research and the RMarkdown [@REF_NEEDED] file used to run and analyse the experiments [@REF_NEEDED].

@@

The individual simulation (`run_simulation()` function) is the working horse in this package. In this function, the ODEs are solved using the function `ode` in the package deSolve package [@Soetaert2010]. The `run_simulation()` function needs only one argument - an object as created by the function `new_runsim_parameter()`. The parameter object contains among other parameter the `strain_parameter` which can be created by the function `new_strain_parameter()`. For a detailed description of the parameter, their meaning and how they are created and can be set and changed please see the User Guide which accompanies the package or is available at [User Guide](@LINK_NEEDED)

After the parameter object has been defined, it can be used in the `run_simulation()` function. The function returns an object which is identical to the parameter object, except of an additional slot containing the results. This design produces a fully reproducible object as it can be used as a parameter object to be fed back into the `run_simulation()` function to run the simulation again from the parameter used to generate the results before.

@@

### Finding a Steady State of the model
\textcolor{red}{@Owen: can you please fill in the references to these mathods? Thanks.}

There are two methods for finding steady states implemented in this package. The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the *Replication method*) \textcolor{red}{{>>@Owen - does this approach has a name, and do you have a reference where it has been used or described?<<}. The second runs only two simulations, with step-wise and slowly temporally increasing or decreasing oxygen diffusivities and recorded of state just before change to a new oxygen diffusivity (let us term this the *Temporal method*)\textcolor{red}{{>>@Owen - Has this been used in other publications? I do not have a clue what I could search for to find out...<<}.

#### Replication Method
The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation. As the multithreading uses the package function `mclapply()` from the package `parallel` [@RCoreTeam2022], the multthreading only works on Linux and Mac. It is planned to move to `parLapply()` [@RCoreTeam2022] in a future release.

#### Temporal Method
The temporal method involves two simulations for a particular system configuration (parameter set). In one simulation the oxygen diffusivity is *increased* in a step-wise fashion. In the other it is *decreased* in a step-wise fashion. That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set. Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step \textcolor{red}{{>>@Owen: is there a ref for this approach?<<}}.

This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()`. It is planned for a later release, to run these two simulations in parallel.

@@

### Extracting Stability Measures
From the raw results returned by these `run_...()` functions, the stability measures can be extracted by using the function `get_stability_measures()`. These measures include non-linearity and hysteresis measures, of the response of the simulated system to environmental change \textcolor{red}{{>>@Owen: is there a paper, so that we can say "as e.g. used in REF?<<}}.

@@

### Regime shifts during temporal environmental change
\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}

@@

### The extent of hysteresis depends on community composition
\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}

@@

### Effects of functional diversity on regime shifts
\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}

@@

\textcolor{red}{{>>@Owen: More details from the paper. I will approach Romana when the first draft is ready so that she can put something in, onless you will add a few sentences here.<<}}

@@

\textcolor{red}{{>>@Owen: Could you put a few sentence in highlighting the impact this package could have considering the importance of possible research questions?<<}}

# Conclusions
> {>>Set out the conclusion of this original software publication.<<}

This package makes an important paper \textcolor{red}{{>>@Owen: is this correct?<<}} [@Bush2017] fully reproducible without having to re-implement the code. By doing so, it opens the door for further investigations based on the system as done in @REF_NEEDED and, consequently, increases the reach (and ultimaly) the impact of the original paper by @Bush2017. This demonstrates again [@REF_NEEDED] the importance of publishing not only the data but also the code used in a publication.

\textcolor{red}{{>>@Owen: could you write a short conclusion paragraph about the ecological relevance of `microxanox`?<<}}

In addition to the ecological relevance and importance of this package as discussed above, it can be used as a template for the implementation of other different models. By doing so, other models can profit from the reproducibility aspects as well as the flexibility as implemented in this package.

### 2022-05-06 inst/manuscript/manuscript_microxanox.tex
Main Changes done

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Microxanox - an R package for simulating an MICRobial ecosystem that can occupy OXic or ANOXic states.},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

@@

  \title{Microxanox - an R package for simulating an \(MICR\)obial
ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
    \author[University of Zürich]{Rainer M Krug%
  %
  \fnref{Corresponding Author}}
   \ead{Rainer.Krug@uzh.ch; Rainer@krugs.de} 
    \author[University of Zürich]{Owen L. Petchey}
   \ead{Owen.Petchey@ieu.uzh.ch} 
      \affiliation[University of Zürich]{Department of Evolutionary
Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{Corresponding Author}
    \fntext[2]{Equal contribution}

  \begin{abstract}
  Ca. 100 words.
  \end{abstract}
    \begin{keyword}
    metadata quality; data curation; archival; long term storage; R
package; \sep 
    metadata quality; data curation; archival; long term storage; R
package;
  \end{keyword}

@@

\hypertarget{todo}{%
\section{\texorpdfstring{\color{red}TODO}{TODO}}\label{todo}}

\begin{itemize}
\tightlist
\item
  \color{red}\textbf{update links}
\item
  \color{red}\textbf{update cross references} \pagebreak
\end{itemize}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0596}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4106}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5298}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule
\endhead
C1 & Current code version & v0.9.0 \\
C2 & Permanent link to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & CC BY 4.0 \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used & R \\
C7 & Compilation requirements, operating environments & R (\textgreater=
4.1.0), ADD PACKAGES \\
C8 & If available Link to developer documentation/manual & \\
C9 & Support email for questions & Rainer.Krug@uzh.ch;
Rainer@krugs.de \\
\bottomrule
\end{longtable}

@@

Many ecosystems are exposed to gradual changes of environmental
variables, to which the responses are not always as gradual as the
change of the environmental variable. One example where the gradual
change of a single environmental variable is causing an abrupt change of
the system is the switch from an aerobic to anaerobic system. This
system has been investigated by \citet{Bush2017}. We wanted to take this
investigation one step further, and look at the role an increased
biodiversity plays in these dynamics \citep{REF_NEEDED}. For this
purpose, we developed this package.

@@

The aims of the \texttt{microxanox} package are twofold. Firstly, the
package aims at reproducing the results shown by \citep{Bush2017}, which
is accomplished in the \href{LINK\%20NEEDED}{vignette Partial
reproduction of Bush et al}. Secondly, to take these results one step
further, it includes new functionality to address our research question
as published in \citep{REF_NEEDED}.

@@

The framework used when writing this package aims at reproducibility of
the results. It builds on the following main considerations:

@@

The point that the results object contains all parameter needed to run
the simulation, promotes reproducibility and makes incremental changes
of individual parameters and re-running the simulations much easier.

@@

\begin{verbatim}
#> [1] "./figures/simflow.pdf"
\end{verbatim}

@@

The individual simulation (\texttt{run\_simulation()} function) is the
working horse in this package. In this function, the ODEs are solved.
The function needs only one argument - an object as created by the
function \texttt{new\_runsim\_parameter()}. One parameter of this object
is the \texttt{strain\_parameter} which can be created by the function
\texttt{new\_strain\_parameter()}. For a detailed description of the
parameter, their meaning and how they are created please see the User
Guide which accompanies the package or is available at
\href{@LINK_NEEDED}{User Guide}

After the parameter object has been defined, it can be used in the
\texttt{run\_simulation()} function. The function returns an object
which is identical to the parameter, except of an additional slot
containing the results. This design produces a fully reproducible object
as it can be used instead of a parameter object to be fed back into the
\texttt{run\_simulation()} function to run the simulation again from the
parameter used to generate the results from.

@@

There are two methods for finding steady states. The first runs a
separate simulation for each combination of starting conditions and
oxygen diffusivity (let us term this the \emph{Replication method}). The
second runs only two simulations, with step-wise and slowly temporally
increasing or decreasing oxygen diffusivities and recorded of state just
before change to a new oxygen diffusivity (let us term this the
\emph{Temporal method}).

@@

The replication method is implementad in the function
\texttt{run\_replication\_ssfind()} which takes a parameter object as
returned by the function \texttt{new\_replication\_ssfind\_parameter()}
and the number of cores for multithreading the simulation.

@@

The temporal method involves two simulations for a particular system
configuration (parameter set). In one simulation the oxygen diffusivity
is \emph{increased} in a step-wise fashion. In the other it is
\emph{decreased} in a step-wise fashion. That is, oxygen diffusivity is
held at a constant level for long enough for steady state to be reach,
that state is recorded, and then a slightly higher (or lower) oxygen
diffusivity value is set. Hence, at that time point, the system is
effectively started with initial conditions that are the state of the
system in the previous time step.

This is implemented in the function \texttt{run\_temporal\_ssfind()},
which takes a parameter object as created by the function
\texttt{new\_temporal\_ssfind\_parameter()}.

@@

From the raw results returned by these \texttt{run\_...()} functions,
the stability measures can be extracted by using the function
\texttt{get\_stability\_measures()}. These measures include
non-linearity and hysteresis measures, of the response of the simulated
system to environmental change.

@@

We will now show two use cases which illustrate applications of the
package. All use cases can be used as starting points for other
investigations of the behavior of the system.

@@

@@

Ass all objects were saved, results can be easily reproduced and
parameter can be changed to assess the impact of these changes (e.g.~to
conduc a sensitivity analysis).

@@

As discussed in the paper \citep{REF_NEEDED}, the role biodiversity
plays in abrupt regime shifts based on gradual changing environmental
parameter is not well understood. This model (as part of the package)
has been used to investigate these dynamics and the results are
available in \citet{REF_NEEDED}.

{\{\textgreater\textgreater More details from the
paper.\textless\textless\} \{\textgreater\textgreater Romana will
provide two or three sentences\textless\textless\}}

@@

@@

Set out the conclusion of this original software publication.

>>>>

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

@@

  \title{Microxanox: an R package for simulating an aquatic
\(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
    \author[University of Zürich]{Rainer M Krug%
  %
  \fnref{Corresponding Author}}
   \ead{Rainer.Krug@uzh.ch; Rainer@krugs.de} 
    \author[University of Zürich]{Owen L. Petchey}
   \ead{Owen.Petchey@ieu.uzh.ch} 
      \affiliation[University of Zürich]{Department of Evolutionary
Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
    \cortext[cor1]{Corresponding author}
    \fntext[1]{Corresponding Author}
    \fntext[2]{Equal contribution}

  \begin{abstract}
  Ca. 100 words.
  \end{abstract}
    \begin{keyword}
    reproducibility; regime shift; \sep 
    reproducibility; regime shift;
  \end{keyword}

@@

\hypertarget{section}{%
\section{\texorpdfstring{\textcolor{red}{TODO}}{}}\label{section}}

\begin{itemize}
\tightlist
\item
  \textbf{\textcolor{red}{update links}}
\item
  \textbf{\textcolor{red}{update cross references}}
\item
  \textbf{\textcolor{red}{check keywords}}
\item
  \textbf{\textcolor{red}{add more references}}
\item
  \textbf{\textcolor{red}{expand on Impact and Conclusion section}}
  \pagebreak
\end{itemize}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule
\endhead
C1 & Current code version & v0.9.0 \\
C2 & Permanent link to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & CC BY 4.0 \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used &
\href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments &
\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
& & \\
& & \\
& & \\
& & \\
& & \\
C8 & If available Link to developer documentation/manual &
\textcolor{red}{TO BE ADDED} \\
C9 & Support email for questions & Rainer.Krug@uzh.ch;
Rainer@krugs.de \\
\bottomrule
\end{longtable}

@@

Many ecosystems are exposed to gradual changes of environmental
variables, to which the responses are not always as gradual as the
change of the environmental variable \citep{REF_NEEDED}. One example
where the gradual change of a single environmental variable is causing
an abrupt change of the system is the switch from an aerobic to
anaerobic system. This system has been investigated by \citet{Bush2017}.
We wanted to take this investigation one step further, and look at the
role an increased biodiversity plays in these dynamics
\citep{REF_NEEDED}. For this purpose, we developed this package.

@@

The aims of the \texttt{microxanox} package are twofold: firstly, to
reproduce the results shown by \citep{Bush2017}, which is accomplished
in the \href{LINK\%20NEEDED}{vignette Partial reproduction of Bush et
al} and secondly, to take these results one step further, to includes
new functionality to address our research question as published in
\citep{REF_NEEDED}.

@@

The framework used when writing this package aims at reproducibility of
the results (see \citep{REF_NEEDED} for examples). It builds on the
following main considerations:

@@

The point that the results object contains all parameter needed to run
the simulation, promotes reproducibility and makes incremental changes
of individual parameters and re-running the simulations much easier, as
can be seen in our research and the RMarkdown \citep{REF_NEEDED} file
used to run and analyse the experiments \citep{REF_NEEDED}.

@@

@@

The individual simulation (\texttt{run\_simulation()} function) is the
working horse in this package. In this function, the ODEs are solved
using the function \texttt{ode} in the package deSolve package
\citep{Soetaert2010}. The \texttt{run\_simulation()} function needs only
one argument - an object as created by the function
\texttt{new\_runsim\_parameter()}. The parameter object contains among
other parameter the \texttt{strain\_parameter} which can be created by
the function \texttt{new\_strain\_parameter()}. For a detailed
description of the parameter, their meaning and how they are created and
can be set and changed please see the User Guide which accompanies the
package or is available at \href{@LINK_NEEDED}{User Guide}

After the parameter object has been defined, it can be used in the
\texttt{run\_simulation()} function. The function returns an object
which is identical to the parameter object, except of an additional slot
containing the results. This design produces a fully reproducible object
as it can be used as a parameter object to be fed back into the
\texttt{run\_simulation()} function to run the simulation again from the
parameter used to generate the results before.

@@

\textcolor{red}{@Owen: can you please fill in the references to these mathods? Thanks.}

There are two methods for finding steady states implemented in this
package. The first runs a separate simulation for each combination of
starting conditions and oxygen diffusivity (let us term this the
\emph{Replication method})
\textbackslash textcolor\{red\}\{\{\textgreater\textgreater{}\citet{Owen}
- does this approach has a name, and do you have a reference where it
has been used or described?\textless\textless\}. The second runs only
two simulations, with step-wise and slowly temporally increasing or
decreasing oxygen diffusivities and recorded of state just before change
to a new oxygen diffusivity (let us term this the \emph{Temporal
method})\textbackslash textcolor\{red\}\{\{\textgreater\textgreater{}\citet{Owen}
- Has this been used in other publications? I do not have a clue what I
could search for to find out\ldots\textless\textless\}.

@@

The replication method is implementad in the function
\texttt{run\_replication\_ssfind()} which takes a parameter object as
returned by the function \texttt{new\_replication\_ssfind\_parameter()}
and the number of cores for multithreading the simulation. As the
multithreading uses the package function \texttt{mclapply()} from the
package \texttt{parallel} \citep{RCoreTeam2022}, the multthreading only
works on Linux and Mac. It is planned to move to \texttt{parLapply()}
\citep{RCoreTeam2022} in a future release.

@@

The temporal method involves two simulations for a particular system
configuration (parameter set). In one simulation the oxygen diffusivity
is \emph{increased} in a step-wise fashion. In the other it is
\emph{decreased} in a step-wise fashion. That is, oxygen diffusivity is
held at a constant level for long enough for steady state to be reach,
that state is recorded, and then a slightly higher (or lower) oxygen
diffusivity value is set. Hence, at that time point, the system is
effectively started with initial conditions that are the state of the
system in the previous time step
\textcolor{red}{{>>@Owen: is there a ref for this approach?<<}}.

This is implemented in the function \texttt{run\_temporal\_ssfind()},
which takes a parameter object as created by the function
\texttt{new\_temporal\_ssfind\_parameter()}. It is planned for a later
release, to run these two simulations in parallel.

@@

From the raw results returned by these \texttt{run\_...()} functions,
the stability measures can be extracted by using the function
\texttt{get\_stability\_measures()}. These measures include
non-linearity and hysteresis measures, of the response of the simulated
system to environmental change
\textcolor{red}{{>>@Owen: is there a paper, so that we can say "as e.g. used in REF?<<}}.

@@

We will now show three use cases which illustrate applications of the
package. All use cases can be used as starting points for other
investigations of the behavior of the system.

@@

\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}

@@

As all objects were saved, results can be easily reproduced and
parameter can be changed to assess the impact of these changes (e.g.~to
conduct a sensitivity analysis).

@@

\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}

As discussed in the paper \citep{REF_NEEDED}, the role biodiversity
plays in abrupt regime shifts based on gradual changing environmental
parameter is not well understood. This model (as part of the package)
has been used to investigate these dynamics and the results are
available in with the DOI DOI NEEDED \citet{REF_NEEDED}.

\textcolor{red}{{>>@Owen: More details from the paper. I will approach Romana when the first draft is ready so that she can put something in, onless you will add a few sentences here.<<}}

@@

\textcolor{red}{{>>@Owen: Could you put a few sentence in highlighting the impact this package could have considering the importance of possible research questions?<<}}

@@

\begin{quote}
\{\textgreater\textgreater Set out the conclusion of this original
software publication.\textless\textless\}
\end{quote}

This package makes an important paper
\textcolor{red}{{>>@Owen: is this correct?<<}} \citep{Bush2017} fully
reproducible without having to re-implement the code. By doing so, it
opens the door for further investigations based on the system as done in
\citet{REF_NEEDED} and, consequently, increases the reach (and ultimaly)
the impact of the original paper by \citet{Bush2017}. This demonstrates
again \citep{REF_NEEDED} the importance of publishing not only the data
but also the code used in a publication.

\textcolor{red}{{>>@Owen: could you write a short conclusion paragraph about the ecological relevance of `microxanox`?<<}}

In addition to the ecological relevance and importance of this package
as discussed above, it can be used as a template for the implementation
of other different models. By doing so, other models can profit from the
reproducibility aspects as well as the flexibility as implemented in
this package.
